### 2024-12-24 2024-5-13/sections/cefa.tex
complete complexity analysis

The semantics of CEFA is defined as follows. Let $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$ be a CEFA. 
A \emph{run} of $\aut$ on a string $w = a_1 \cdots a_n$ is a sequence $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ such that $q_0 \in I$ and $q_{i-1} \xrightarrow[\myvec{v_i}]{a_i} q_i$ for each $i \in [n]$. A run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$ is \emph{accepting} if $q_n \in F$ and $\alpha(\myvec{v'}/R)$ is true, where $\myvec{v'} = \sum \limits_{j \in [n]} \myvec{v_j}$. The string $w$ is $accepted$ by the CEFA if and only if there is an accepting run of $w$. The vector $\myvec{v'}= \sum \limits_{j \in [n]} \myvec{v_j}$ is called the \emph{cost} of an accepting run $q_0 \xrightarrow[\myvec{v_1}]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v_n}]{a_n} q_n$.Note that the values of all registers are initiated to zero and updated to $\sum \limits_{j \in [n]} \myvec{v_j}$ after all the transitions in the run are executed. We use $\myvec{v'} \in \aut(w)$ to denote the fact that there is an accepting run of $\aut$ on $w$ whose cost is $\myvec{v'}$.  
We define the semantics of a CEFA $\aut$, denoted by $\Lang(\aut)$, as $\{(w; \myvec{v'}) \mid  \myvec{v'} \in \aut(w)\}$.  In particular, if $I \cap F \neq \emptyset$ and $\alpha[\myvec{0}/R]$ is true, then $(\varepsilon; \myvec{0}) \in \Lang(\aut)$. Moreover, we define the \emph{output} of a CEFA $\aut$, denoted by $\cefaout(\aut)$, as $\{\myvec{v'} \mid  \exists w.\ \myvec{v'} \in \aut(w)\}$.

@@

Let $\aut_1 = (R_1, Q_1, \Sigma, \delta_1, q_{1,0}, F_1, \alpha_1)$ and  $\aut_2 = (R_2, Q_2, \Sigma, \delta_2, q_{2,0}, F_2, \alpha_2)$ be two CEFA that share the alphabet. 
Moreover, suppose that $R_1 \cap R_2 = \emptyset$ and $Q_1 \cap Q_2 = \emptyset$.

The \emph{union} of $\aut_1$ and $\aut_2$ is denoted by $\aut_1 \cup \aut_2$. Two fresh auxiliary registers say $r'_1,r'_2 \not \in R_1 \cup R_2$, are introduced so that the accepting condition knows whether a run is from $\aut_1$ (or $\aut_2$). Specifically, $\aut_1 \cup \aut_2 = (R', Q', \Sigma, \delta', I', F', \alpha')$ where 
\begin{itemize}
\item $R' = R_1 \cup R_2 \cup \{r'_1, r'_2\}$, $Q' = Q_1 \cup Q_2 \cup \{q'_0\}$ with $q'_0 \not \in Q_1 \cup Q_2$, $I' = \{q'_0\}$, 
%
\item $\delta'$ is the union of four transitions sets:
\begin{itemize}
  \item $\{(q'_0, a, q'_1, (\myvec{v_1}, \myvec{0}, 1,0)) \mid \exists q_1 \in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q'_0, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 1)) \mid \exists q_2 \in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
  \item $\{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0}, 0, 0)) \mid q_1\not\in I_1.\ (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\}$
  \item $\{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2}, 0, 0)) \mid q_2\not\in I_2.\ (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\}$
\end{itemize}
 where $(\myvec{v_1}, \myvec{0}, 1,0)$ is a vector that updates $R_1$ by $\myvec{v_1}$, updates $R_2$ by $\myvec{0}$, and updates $r'_1,r'_2$ by $1,0$ respectively. Similarly for $(\myvec{0}, \myvec{v_2}, 0, 1)$, and so on.

@@

The \emph{intersection} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \cap \aut_2 = (R', Q', \Sigma, \delta', I', F', \alpha')$, is defined in the sequel. 
\begin{itemize}
\item $R' = R_1 \cup R_2$, $Q' = Q_1 \times Q_2$, $I' = I_1 \times I_2$, $F' = F_1 \times F_2$, $\alpha' = \alpha_1 \wedge \alpha_2$, 
\item $\delta'$ comprises the tuples $((q_1, q_2), a, (q'_1, q'_2), (\myvec{v_1}, \myvec{v_2}))$ such that $(q_1,a, q'_1, \myvec{v_1}) \in \delta_1$ and $(q_2,a, q'_2, \myvec{v_2}) \in \delta_2$.
\end{itemize}
From the construction, 
$$\Lang(\aut_1 \cap \aut_2) = \{(w; \myvec{v}) \mid (w; \prj_{R_1}(\myvec{v})) \in \Lang(\aut_1) \mbox{ and } (w; \prj_{R_2}(\myvec{v})) \in \Lang(\aut_2)\}.$$
Intuitively, $\aut_1 \cap \aut_2$ accepts the words that are accepted by both $\aut_1$ and $\aut_2$ and outputs the costs of $\aut_1$ and $\aut_2$.

The \emph{concatenation} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \concat \aut_2$, is defined similarly as that of NFA, that is, a tuple $(Q', \Sigma, \delta', I', F', \alpha')$, where $Q' = Q_1 \cup Q_2$, $I' = I_1$, $\alpha' = \alpha_1 \wedge \alpha_2$, $\delta' = \{(q_1, a, q'_1, (\myvec{v_1}, \myvec{0})) \mid (q_1, a, q'_1, \myvec{v_1}) \in \delta_1\} \cup \{(q_2, a, q'_2, (\myvec{0}, \myvec{v_2})) \mid (q_2, a, q'_2, \myvec{v_2}) \in \delta_2\} \cup \{(q_1, a, q_2, (\myvec{0}, \vec{v_2})) \mid q_1 \in F_1, \exists q' \in I_2.\ (q', a, q_2, \vec{v_2}) \in \delta_2\}$, moreover, if $I_2 \cap F_2 \neq \emptyset$, then $F'= F_1 \cup F_2$, otherwise, $F'= F_2$. From the construction, 
$$\Lang(\aut_1 \concat \aut_2) = \{(w_1w_2; \myvec{v}) \mid (w_1; \prj_{R_1}(\myvec{v})) \in \Lang(\aut_1) \mbox{ and }(w_2; \prj_{R_2}(\myvec{v})) \in \Lang(\aut_2)\}.$$

>>>>

The semantics of CEFA is defined as follows. Let $\aut = (R, Q, \Sigma, \delta, I, F, \alpha)$ be a CEFA. 
A \emph{run} of $\aut$ on a string $w = a_1 \cdots a_n$ is a sequence $q_0 \xrightarrow[\myvec{v}_1]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v}_n]{a_n} q_n$ such that $q_0 \in I$ and $q_{i-1} \xrightarrow[\myvec{v}_i]{a_i} q_i$ for each $i \in [n]$. A run $q_0 \xrightarrow[\myvec{v}_1]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v}_n]{a_n} q_n$ is \emph{accepting} if $q_n \in F$ and $\alpha(\myvec{v'}/R)$ is true, where $\myvec{v'} = \sum \limits_{j \in [n]} \myvec{v}_j$. The string $w$ is $accepted$ by the CEFA if and only if there is an accepting run of $w$. The vector $\myvec{v'}= \sum \limits_{j \in [n]} \myvec{v}_j$ is called the \emph{cost} of an accepting run $q_0 \xrightarrow[\myvec{v}_1]{a_1} q_1 \cdots q_{n-1}\xrightarrow[\myvec{v}_n]{a_n} q_n$.Note that the values of all registers are initiated to zero and updated to $\sum \limits_{j \in [n]} \myvec{v}_j$ after all the transitions in the run are executed. We use $\myvec{v'} \in \aut(w)$ to denote the fact that there is an accepting run of $\aut$ on $w$ whose cost is $\myvec{v'}$.  
We define the semantics of a CEFA $\aut$, denoted by $\Lang(\aut)$, as $\{(w; \myvec{v'}) \mid  \myvec{v'} \in \aut(w)\}$.  In particular, if $I \cap F \neq \emptyset$ and $\alpha[\myvec{0}/R]$ is true, then $(\varepsilon; \myvec{0}) \in \Lang(\aut)$. Moreover, we define the \emph{output} of a CEFA $\aut$, denoted by $\cefaout(\aut)$, as $\{\myvec{v'} \mid  \exists w.\ \myvec{v'} \in \aut(w)\}$.

@@

Let $\aut_1 = (R_1, Q_1, \Sigma, \delta_1, \hdh{I_1}, F_1, \alpha_1)$ and  $\aut_2 = (R_2, Q_2, \Sigma, \delta_2, \hdh{I_2}, F_2, \alpha_2)$ be two CEFA that share the alphabet. 
Moreover, suppose that $R_1 \cap R_2 = \emptyset$ and $Q_1 \cap Q_2 = \emptyset$.

The \emph{union} of $\aut_1$ and $\aut_2$ is denoted by $\aut_1 \cup \aut_2$. Two fresh auxiliary registers say $r'_1,r'_2 \not \in R_1 \cup R_2$, are introduced so that the accepting condition knows whether a run is from $\aut_1$ (or $\aut_2$).\hdh{This is mainly because the same accepting condition is applied to all final states, in other words, the accepting condition does not distinguish between final states.} Specifically, $\aut_1 \cup \aut_2 = (R', Q', \Sigma, \delta', I', F', \alpha')$ where 
\begin{itemize}
\item $R' = R_1 \cup R_2 \cup \{r'_1, r'_2\}$, $Q' = Q_1 \cup Q_2 \cup \{q'_0\}$ with $q'_0 \not \in Q_1 \cup Q_2$, $I' = \{q'_0\}$, 
%
\item $\delta'$ is the union of four transitions sets:
\begin{itemize}
  \item $\{(q'_0, a, q'_1, (\myvec{v}_1, \myvec{0}, 1,0)) \mid \exists q_1 \in I_1.\ (q_1, a, q'_1, \myvec{v}_1) \in \delta_1\}$
  \item $\{(q'_0, a, q'_2, (\myvec{0}, \myvec{v}_2, 0, 1)) \mid \exists q_2 \in I_2.\ (q_2, a, q'_2, \myvec{v}_2) \in \delta_2\}$
  \item $\{(q_1, a, q'_1, (\myvec{v}_1, \myvec{0}, 0, 0)) \mid q_1\not\in I_1.\ (q_1, a, q'_1, \myvec{v}_1) \in \delta_1\}$
  \item $\{(q_2, a, q'_2, (\myvec{0}, \myvec{v}_2, 0, 0)) \mid q_2\not\in I_2.\ (q_2, a, q'_2, \myvec{v}_2) \in \delta_2\}$
\end{itemize}
 where $(\myvec{v}_1, \myvec{0}, 1,0)$ is a vector that updates $R_1$ by $\myvec{v}_1$, updates $R_2$ by $\myvec{0}$, and updates $r'_1,r'_2$ by $1,0$ respectively. Similarly for $(\myvec{0}, \myvec{v}_2, 0, 1)$, and so on.

@@

The \emph{intersection} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \cap \aut_2 = (R', Q', \Sigma, \delta', I', F',$ $ \alpha')$, is defined in the sequel. 
\begin{itemize}
\item $R' = R_1 \cup R_2$, $Q' = Q_1 \times Q_2$, $I' = I_1 \times I_2$, $F' = F_1 \times F_2$, $\alpha' = \alpha_1 \wedge \alpha_2$, 
\item $\delta'$ comprises the tuples $((q_1, q_2), a, (q'_1, q'_2), (\myvec{v}_1, \myvec{v}_2))$ such that $(q_1,a, q'_1, $ $\myvec{v}_1)\in \delta_1$ and $(q_2,a, q'_2, \myvec{v}_2) \in \delta_2$.
\end{itemize}
From the construction, 
$$\Lang(\aut_1 \cap \aut_2) = \{(w; \myvec{v}) \mid (w; \prj_{R_1}(\myvec{v})) \in \Lang(\aut_1) \mbox{ and } (w; \prj_{R_2}(\myvec{v})) \in \Lang(\aut_2)\}.$$
Intuitively, $\aut_1 \cap \aut_2$ accepts the words that are accepted by both $\aut_1$ and $\aut_2$ and outputs the costs of $\aut_1$ and $\aut_2$.

The \emph{concatenation} of $\aut_1$ and $\aut_2$, denoted by $\aut_1 \concat \aut_2$, is defined similarly as that of NFA, that is, a tuple $(\hdh{R'}, Q', \Sigma, \delta', I', F', \alpha')$, where \hdh{$R' = R_1\cup R_2$}, $Q' = Q_1 \cup Q_2$, $I' = I_1$, $\alpha' = \alpha_1 \wedge \alpha_2$, $\delta' = \{(q_1, a, q'_1, (\myvec{v}_1, \myvec{0})) \mid (q_1, a, q'_1, \myvec{v}_1) \in \delta_1\} \cup \{(q_2, a, q'_2, (\myvec{0}, \myvec{v}_2)) \mid (q_2, a, q'_2, \myvec{v}_2) \in \delta_2\} \cup \{(q_1, a, q_2, (\myvec{0}, \vec{v_2})) \mid q_1 \in F_1, \exists q' \in I_2.\ (q', a, q_2, \vec{v_2}) \in \delta_2\}$, moreover, if $I_2 \cap F_2 \neq \emptyset$, then $F'= F_1 \cup F_2$, otherwise, $F'= F_2$. From the construction, 
$$\Lang(\aut_1 \concat \aut_2) = \{(w_1w_2; \myvec{v}) \mid (w_1; \prj_{R_1}(\myvec{v})) \in \Lang(\aut_1) \mbox{ and }(w_2; \prj_{R_2}(\myvec{v})) \in \Lang(\aut_2)\}.$$

### 2024-12-24 2024-5-13/sections/heuristic.tex
complete complexity analysis

\item We estimate the sizes of the two CEFAs $\aut_{\ufld_1(e)}, \aut_{\ufld_2(e)}$ constructed from $\ufld_1(e), \ufld_2(e)$ respectively as pairs $(\snum_1(e), \rnum_1(e))$ and $(\snum_2(e), \rnum_2(e))$, where $\snum_1(e)$ and $\rnum_1(e)$ are the estimates of the number of states and registers in $\aut_{\ufld_1(e)}$, similarly for $\snum_2(e)$ and $\rnum_2(e)$. 
%
\item From the construction in Section~\ref{sec:algorithm}, 
two NFAs $\cU_{\aut_{\ufld_\hdh{1}(e)}}, \cU_{\aut_{\ufld_2(e)}}$ are constructed from $\aut_{\ufld_1(e)}$ and $\aut_{\ufld_2(e)}$ by dropping the accepting conditions and ignoring the characters. As a result, the alphabets of $\cU_{\aut_{\ufld_\hdh{1}(e)}}, \cU_{\aut_{\ufld_2(e)}}$ are the set of integer vectors \sout{$\myvec{v}$} occurring in the transitions of $\aut_{\ufld_1(e)}, \aut_{\ufld_2(e)}$ respectively. \sout{The integer vectors of  $\aut_{\ufld_1(e)}$ are of form $(\underbrace{0, \cdots, 0}_{i-1}, 1, \underbrace{0, \cdots, 0}_{r'-i-1})$, where $i \in [r']$.} \hdh{These integer vectors are of form $(0,\cdots,0,1,0,\cdots, 0)$ whose length is the number of registers.} Consequently, the size of the alphabet of $\cU_{\aut_{\ufld_1(e)}}$ is bounded by $\rnum_1(e)$. Similarly for $\cU_{\aut_{\ufld_\hdh{2}(e)}}$. 
Finally, we estimate the sizes (number of transitions) of the NFAs $\cU_{\aut_{\ufld_1(e)}}, \cU_{\aut_{\ufld_2(e)}}$ as $sz_1 = (\snum_1(e))^2 \cdot \rnum_1(e)$ and $sz_2 = (\snum_2(e))^2 \cdot \rnum_2(e)$ respectively.
%
\item If $sz_1 \le sz_2$, then we choose to unfold the counting operators in $e_1$. Otherwise, we choose to unfold the counting operator $^{\{m,n\}}$. 
\end{itemize}

@@

It remains to show how to compute the pairs $(\snum_1(e), \rnum_1(e))$ and $(\snum_2(e), \rnum_2(e))$ from $e=(e_1)^{\{m,n\}}$. Let $e=(e_1)^{\{m,n\}}$ be a regular expression where $e_1$ is a register-representable regular expression. Then we compute $\snum'_1(e_1)$ and $(\snum'_2(e_1), \rnum'_2(e_1))$\hdh{(TODO:Change all the $num'$ to $num$, to avoid ambiguity)} by a structural induction on $e_1$, \hdh{in condition, $\rnum_1(e_1)$ is set to $1$.} \sout{Finally, let $(\snum_1(e), \rnum_1(e)) = (\snum'_1(e_1), 1)$ and $(\snum_2(e), \rnum_2(e)) = (\snum'_2(e_1), \rnum'_2(e_1))$.}
\begin{itemize}
  \item If $e_1 = \emptyset$ or $e_1 = \epsilon$, then $\snum'_1(e_1)  = 0$ and $(\snum'_2(e_1), \rnum'_2(e_1))  = (0, 0)$. 
%
  \item \hdh{If $e_1 = a$, then $\snum'_1(e_1) = 1$ and ($\snum'_2(e_1), \rnum'_2(e_1)) = (n, 0)$.}
 %
  \item If $e_1 = e_2 \concat e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) + \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) + \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3))).
 \end{array}
  $$
%
 \item If $e_1 = e_2 + e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) + \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) + \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3)+1)).
 \end{array}
  $$
 (Recall that the construction of $\aut_1 \cup \aut_2$ introduces one additional register.)
%
 \item If $e_1 = e_2 \cap e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) \times \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) \times \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3))).
 \end{array}
  $$
%
 \item If $e_1 = \overline{e_2}$, then $\snum'_1(e_1) = 2^{\snum'_1(e_2)}$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \cdot 2^{\snum'_2(e_2)}, 0).
% \end{array}
  $
 (The exponential blow-up is explained by the fact that the determinization is needed for the complementation.) 
%
%  \item If $\regex = \overline{\regex_1}$, then $s_\regex = 2^{s_{\regex_1}}$ (since the determinization is needed for the complement) and $r_\regex = 0$.
%
 \item If $e_1 = (e_2)^*$, then $\snum'_1(e_1) = \snum'_1(e_2)$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \snum'_2(e_2), 0).
% \end{array}
  $
%  \item If $\regex = \regex_1^*$, then $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 0$.
%  \item $\regex = \regex_1 \backslash \regex_2$ can be transferred to $\regex = \regex_1 \cap \overline{\regex_2}$ and then the size can be computed by the formula of $\cap$ and $\bar{}$.
 \item If $e_1 = (e_2)^{\{m', n'\}}$ or $e_1 = (e_2)^{\{n', \infty\}}$, then $\snum'_1(e_1) = n' \snum'_1(e_2)$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \snum'_2(e_2), n).
% \end{array}
  $
%  \item For $\regex = \regex_1^{\{m, n\}}$ or $\regex = \regex_1^{\{n,\infty\}}$, $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 1$.
\end{itemize}

@@

Without loss of generality, we can assume that there are no transitions out of $F$ in $\aut_{e_1}$. If this is not the case, then a new state $q_f$ can be introduced to transform $\aut_{e_1}$ into a CEFA satisfying this constraint, where $q_f$ is the unique accepting state.\footnote{\hdh{Note that if any initial states are in $F$, a new state $q_I$ without transitions out can be introduced and add to $F$, to accept empty string. Then we can remove all original initial states out of $F$.}} 
Then we define $\cB= (R \cup \{r'\}, Q \cup \{q_s\}, \Sigma, \delta', I, Q \cup \{q_s\}, \alpha')$, where
\begin{itemize}
  \item $r'$ is a new register not in $R$, 
  \item $q_s$ is a new state not in $Q$,
  \item $\delta'$ comprises the following transitions, 
  \begin{itemize}
    \item $(q, a, q', (\vec{v}, 0))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \not \in F$,
%
    \item $(q, a, q_s,(\vec{0}, 0))$ such that there do \emph{not} exist $a$-labeled transitions out of $q$, that is, for every $q'$ and $\vec{v}$, we have $(q, a, q', \vec{v}) \not\in \delta$,
%
    \item $(q, a, q', (\vec{v}, 1))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \in F$,
%
    \item $(q_s, a, q_s, (\vec{0}, 0))$ such that $a \in \Sigma$, 
  \end{itemize}
  \item $\alpha'$ is defined as $r' = 1 \rightarrow \neg \alpha$.
\end{itemize}
Intuitively, $q_s$ is a sink state that deals with the situations where there are no runs, and $r'$ is a register newly introduced so that the different cases of the non-accepting runs of $\aut_{e_1}$ can be captured by a single accepting condition $\alpha'$.

@@

\hdh{\paragraph*{Combination of under- and over-approximation}
Since these two types of approximation influence each other, we separate them apart and run them parallelly.
}

@@

\hdh{The naive approach of finding satisfied string models is to search the accepted words on CEFAs based on the integer models of vectors, which causes an exponential blow-up of searching space. Specifically, } recall that in Section~\ref{sec:algorithm}, the satisfiability of the original RECL constraint is reduced to the satisfiability of the existential LIA formula
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right].
\end{array}
\]
Let us use $\theta$ to denote this formula. 
Then we utilize some SMT solver, say Z3, to solve the satisfiability of $\theta$. If Z3 reports ``sat'', then it also returns a model, namely, an assignment $\amodel$ of natural numbers to the free integer variables in $\theta$. Then we utilize $\amodel$ to generate a model $\amodel'$ of the original RECL constraint $\varphi$, i.e., an assignment of strings to the string variables so that $\varphi$ is evaluated to true under $\amodel'$.

>>>>

\item We estimate the sizes of the two CEFAs $\aut_{\ufld_1(e)}, \aut_{\ufld_2(e)}$ constructed from $\ufld_1(e), \ufld_2(e)$ respectively as pairs $(\snum_1(e), \rnum_1(e))$ and $(\snum_2(e), \rnum_2(e))$, where $\snum_1(e)$ and $\rnum_1(e)$ are the estimates of the number of states and registers in $\aut_{\ufld_1(e)}$, similarly for $\snum_2(e)$ and $\rnum_2(e)$. 
%
\item From the construction in Section~\ref{sec:algorithm}, 
two NFAs $\cU_{\aut_{\ufld_\hdh{1}(e)}}, \cU_{\aut_{\ufld_2(e)}}$ are constructed from $\aut_{\ufld_1(e)}$ and $\aut_{\ufld_2(e)}$ by dropping the accepting conditions and ignoring the characters. As a result, the alphabets of $\cU_{\aut_{\ufld_\hdh{1}(e)}}, \cU_{\aut_{\ufld_2(e)}}$ are the set of integer vectors \sout{$\myvec{v}$} occurring in the transitions of $\aut_{\ufld_1(e)}, \aut_{\ufld_2(e)}$ respectively. \sout{The integer vectors of  $\aut_{\ufld_1(e)}$ are of form $(\underbrace{0, \cdots, 0}_{i-1}, 1, \underbrace{0, \cdots, 0}_{r'-i-1})$, where $i \in [r']$.} \hdh{The integer vectors of  $\aut_{\ufld_1(e)}$ are of the form $(\underbrace{0, \cdots, 0}_{i-1}, 1, \underbrace{0, \cdots, 0}_{\rnum_1(e)-i-1})$, where $i \in [\rnum_1(e)]$.} Consequently, the size of the alphabet of $\cU_{\aut_{\ufld_1(e)}}$ is bounded by $\rnum_1(e)$. Similarly for $\cU_{\aut_{\ufld_\hdh{2}(e)}}$. 
Finally, we estimate the sizes (number of transitions) of the NFAs $\cU_{\aut_{\ufld_1(e)}}, \cU_{\aut_{\ufld_2(e)}}$ as $sz_1 = (\snum_1(e))^2 \cdot \rnum_1(e)$ and $sz_2 = (\snum_2(e))^2 \cdot \rnum_2(e)$ respectively.
%
\item If $sz_1 \le sz_2$, then we choose to unfold the counting operators in $e_1$. Otherwise, we choose to unfold the counting operator $^{\{m,n\}}$. 
\end{itemize}

@@

It remains to show how to compute the pairs $(\snum_1(e), \rnum_1(e))$ and $(\snum_2(e), \rnum_2(e))$ from $e=(e_1)^{\{m,n\}}$. Let $e=(e_1)^{\{m,n\}}$ be a regular expression where $e_1$ is a register-representable regular expression. Then we compute $\snum'_1(e_1)$ and $(\snum'_2(e_1), \rnum'_2(e_1))$. Finally, let $(\snum_1(e), \rnum_1(e)) = (\snum'_1(e_1), 1)$ and $(\snum_2(e), \rnum_2(e)) = (\snum'_2(e_1), \rnum'_2(e_1))$.
\begin{itemize}
  \item If $e_1 = \emptyset$ or $e_1 = \epsilon$, then $\snum'_1(e_1)  = 0$ and $(\snum'_2(e_1), \rnum'_2(e_1))  = (0, 0)$. 
%
  \item \hdh{If $e_1 = a$, then $\snum'_1(e_1) = 1$ and ($\snum'_2(e_1), \rnum'_2(e_1)) = (n, 0)$.}
 %
  \item If $e_1 = e_2 \concat e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) + \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) + \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3))).
 \end{array}
  $$
%
 \item If $e_1 = e_2 + e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) + \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) + \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3)+1)).
 \end{array}
  $$
 (Recall that the construction of $\aut_1 \cup \aut_2$ introduces one additional register.)
%
 \item If $e_1 = e_2 \cap e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) \times \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) \times \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3))).
 \end{array}
  $$
%
 \item If $e_1 = \overline{e_2}$, then $\snum'_1(e_1) = 2^{\snum'_1(e_2)}$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \cdot 2^{\snum'_2(e_2)}, 0).
% \end{array}
  $
 (The exponential blow-up is explained by the fact that the determinization is needed for the complementation.) 
%
%  \item If $\regex = \overline{\regex_1}$, then $s_\regex = 2^{s_{\regex_1}}$ (since the determinization is needed for the complement) and $r_\regex = 0$.
%
 \item If $e_1 = (e_2)^*$, then $\snum'_1(e_1) = \snum'_1(e_2)$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \snum'_2(e_2), 0).
% \end{array}
  $
%  \item If $\regex = \regex_1^*$, then $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 0$.
%  \item $\regex = \regex_1 \backslash \regex_2$ can be transferred to $\regex = \regex_1 \cap \overline{\regex_2}$ and then the size can be computed by the formula of $\cap$ and $\bar{}$.
 \item If $e_1 = (e_2)^{\{m', n'\}}$ or $e_1 = (e_2)^{\{n', \infty\}}$, then $\snum'_1(e_1) = n' \snum'_1(e_2)$ and 
%
  $
%  \begin{array}{l }
 (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \snum'_2(e_2), n).
% \end{array}
  $
%  \item For $\regex = \regex_1^{\{m, n\}}$ or $\regex = \regex_1^{\{n,\infty\}}$, $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 1$.
\end{itemize}

@@

Without loss of generality, we can assume that there are no transitions out of $F$ in $\aut_{e_1}$. If this is not the case, then a new state $q_f$ can be introduced to transform $\aut_{e_1}$ into a CEFA satisfying this constraint, where $q_f$ is the unique accepting state. 
Then we define $\cB= (R \cup \{r'\}, Q \cup \{q_s\}, \Sigma, \delta', I, Q \cup \{q_s\}, \alpha')$, where
\begin{itemize}
  \item $r'$ is a new register not in $R$, 
  \item $q_s$ is a new state not in $Q$,
  \item $\delta'$ comprises the following transitions, 
  \begin{itemize}
    \item $(q, a, q', (\vec{v}, 0))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \not \in F$,
%
    \item $(q, a, q_s,(\vec{0}, 0))$ such that there do \emph{not} exist $a$-labeled transitions out of $q$, that is, for every $q'$ and $\vec{v}$, we have $(q, a, q', \vec{v}) \not\in \delta$,
%
    \item $(q, a, q', (\vec{v}, 1))$ such that $(q, a, q', \vec{v}) \in \delta$ and $q' \in F$,
%
    \item $(q_s, a, q_s, (\vec{0}, 0))$ such that $a \in \Sigma$, 
  \end{itemize}
  \item $\alpha'$ is defined as $r' = 1 \rightarrow \neg \alpha$.
\end{itemize}
Intuitively, $q_s$ is a sink state that deals with the situations where there are no runs, and $r'$ is a register newly introduced so that the different cases of the non-accepting runs of $\aut_{e_1}$ can be captured by a single accepting condition $\alpha'$.

@@

\hdh{\paragraph*{Combination of under- and over-approximation}
 In our configuration, under-approximation is performed first due to its superior computational efficiency, and we proceed with over-approximation if it yields an unsatisfiable result. In cases where over-approximation produces a satisfiable result, we execute the decision procedure wherein all non-register-representable regular expressions are naively unfold. To minimize computational overhead, we apply approximation techniques only in cases involving non-register-representable regular expressions with significant counting bounds—specifically, those exceeding 20 in our experimental configuration.
}

@@

\zhilin{TODO: I had troubles to understand the section about optimized model generation. Maybe stating what the technical problem that it is solving is in a more dumb proof way would help., } recall that in Section~\ref{sec:algorithm}, the satisfiability of the original RECL constraint is reduced to the satisfiability of the existential LIA formula
\[
\begin{array}{l}
\bigwedge \limits_{i \in [n]} \left(\psi_{\cC_i}(\fZ_{\costset_{\cC_i}}) \wedge \alpha_i\left[\sum\limits_{\myvec{v} \in \costset_{\cC_i}} \anivarz_{i, \myvec{v}} \myvec{v} \big{/}R_{\cB_i}  \right] \right) \ \wedge \\
\ \ \varphi \left[ \sum\limits_{\myvec{v} \in \costset_{\cC_1}} \anivarz_{1, \myvec{v}} \myvec{v} \big{/}R_{\cB_1}, \cdots, \sum\limits_{\myvec{v} \in \costset_{\cC_n}} \anivarz_{n, \myvec{v}} \myvec{v} \big{/}R_{\cB_n} \right].
\end{array}
\]
Let us use $\theta$ to denote this formula. 
Then we utilize some SMT solver, say Z3, to solve the satisfiability of $\theta$. If Z3 reports ``sat'', then it also returns a model, namely, an assignment $\amodel$ of natural numbers to the free integer variables in $\theta$. Then we utilize $\amodel$ to generate a model $\amodel'$ of the original RECL constraint $\varphi$, i.e., an assignment of strings to the string variables so that $\varphi$ is evaluated to true under $\amodel'$.
